---
title: "ACE2 Figures"
author: "Mareike"
date: "4/3/2020"
output: html_document
---

Load packages:
```{r}
library(ggtree)
library(ggplot2)
library(ape)
```

Read in full mammal tree from TimeTree:
```{r}
mammal_tree <- read.tree("../Mammalia_species.nwk")
```

Drop tips that are not in species list: (necessary because TimeTree species list upload not working at the time)
```{r}
species_list <- read.delim("../species_list_ACE2.txt", header = FALSE)
species_list$tips <- gsub(' ', '_', species_list$V1)
tips <- as.vector(species_list$tips)

ACE2_tr <- keep.tip(mammal_tree, tips)
```

Read in the table with data on covid susceptibility to color the branches of the phylogeny:
```{r}
risk_df <- read.delim("../Metadata/covid-risk.tsv", header = TRUE, sep = "\t")

#making a separate column with species names without underscores to use as tiplabels for the tree
risk_df$tiplabel <- as.character(gsub("_", " ", risk_df$Species))
```

Detach ape to use ggtree rotate function:
```{r}
detach("package:ape")
```

Order levels of risk variable for figure:
```{r}
risk_df$Risk <- factor(risk_df$Risk, levels = c("known", "predicted", "uncertain", "low", "none"))
```


Plot tree and rotate branches:
```{r}
tr1 <- ggtree(ACE2_tr)
tr2 <- rotate(tr1, 62) %>% rotate(72) %>% rotate(76)

p_tr2 <- tr2 %<+% risk_df + 
  # adding color to the branches was omitted here and instead done in Illustrator, because msaplot throws an error, if included. See below for another option to generate a combo figure for phylogeny and multiple sequence alignment.
  #aes(color = Risk) + 
  scale_color_manual(name = "Covid-19 susceptibility", values = c("#F21A00", "#E58600", "#E8C520", "#82B8B6", "#3B9AB2"), na.translate = TRUE, na.value = "black", guide = "legend") +
  geom_tiplab(color = "black", fontface='italic', hjust = 1, align = TRUE, offset = 30, aes(label=tiplabel))

p_tr2 
```

Add the plot of the multiple sequence alignment (MSA):
```{r}
p_msa <- msaplot(p_tr2, "../Alignments/ACE2_MSA_for_figure.fasta", offset = 33, width = 0.9)
p_msa

#ggsave("/Users/Mareike/Documents/Data/ACE2/Figures/concept_figure.pdf", plot = p_msa, width = 21, height = 10, units = "in", useDingbats=FALSE)
```


Alternate option for generating a figure of a phylogeny with adjacent multiple sequence alignment:

Make tree object:
```{r}
tr1a <- ggtree(ACE2_tr)
tr2a <- rotate(tr1a, 62) %>% rotate(72) %>% rotate(76)

p_tr2a <- tr2a %<+% risk_df + 
  aes(color = Risk) + 
  scale_color_manual(name = "Covid-19 susceptibility", values = c("#F21A00", "#E58600", "#E8C520", "#82B8B6", "#3B9AB2"), na.translate = TRUE, na.value = "black", guide = "legend") +
  geom_tiplab(color = "black", fontface='italic', hjust = 1, align = TRUE, offset = 30, aes(label=tiplabel))

p_tr2a
```

Add the multiple sequence alignment with the ggmsa package:
```{r}
library(Biostrings)
library (ggmsa)
x <- readAAStringSet("../Alignments/ACE2_MSA_for_figure.fasta")
d <- as.dist(stringDist(x, method = "hamming")/width(x)[1])
data = tidy_msa(x)

facet_plot(p = p_tr2a, geom = geom_msa, data = data,  panel = 'msa', font = NULL, color = "Chemistry_AA") + xlim_tree(1)
```

Note: An unfortunate issue with both approaches is that the dotted line created by "align = TRUE" continues below the tip label. Haven't been able to find a way to adjust this yet, but these extra dots can be removed in Illustrator.

### Plotting gene trees for Supplemental Figure

* Read in trees:
```{r}
MrBayes_tree <- read.tree("../GeneTrees/ACE2_MrBayes_tree.newick")
RAxML_tree <- read.tree("../GeneTrees/ACE2_RAxML_tree.newick")
```

* Create dataframe for tip labels:
```{r}
tiplabs <- data.frame(MrBayes_tree$tip.label)
tiplabs$tips <- gsub("_", " ", tiplabs$MrBayes_tree.tip.label)
```

RAxML tree:
```{r}
p_tr_ML1 <- ggtree(RAxML_tree) + geom_treescale() + geom_text2(size = 3.5, aes(subset = !isTip, label=label, hjust = 1.2, vjust = -0.8))
p_tr_ML <- p_tr_ML1 %<+% tiplabs + geom_tiplab(fontface = "italic", aes(label = tips))
p_tr_ML
```

MrBayes tree:
```{r}
p_tr_Bayes1 <- ggtree(MrBayes_tree) + geom_treescale() + geom_text2(size = 3.5, aes(subset = !isTip, label=label, hjust = 1.2, vjust = -0.8))
p_tr_Bayes <- p_tr_Bayes1 %<+% tiplabs + geom_tiplab(fontface = "italic", aes(label = tips))
p_tr_Bayes
```

Combine both trees into one figure:
```{r}
library(ggpubr)

p_genetrees <- ggarrange(p_tr_Bayes, p_tr_ML, ncol = 1, labels = "auto")

#ggsave("/Users/Mareike/Documents/Data/ACE2/Figures/ACE2_GeneTree.pdf", plot = p_genetrees, width = 20, height = 15, units = "in", useDingbats = FALSE)
```
